Simulation Tools for Small Area Estimation: Introducing the \proglang{R}-package \proglang{saeSim}

### 2014-10-31 saeSim.tex
update Paper saeSim

Without knowing anything about the setup defined in \proglang{sim\_base\_lm} we can see that \proglang{setup1} and \proglang{setup2} only differ in the applied sampling scheme. \proglang{sim\_sample} is responsible to control when a function is applied (after the population-level) and \proglang{sample\_number(5)} and \proglang{sample\_fraction(0.05)} define the explicit way of drawing samples. Braking the responsibility of each component into what is applied and when it is applied makes it possible to add new components to any step in the process. The composition of a simulation in that manner will focus on the definition of components and hide control structures. Any function can be passed to \proglang{sim\_sample} which has a \proglang{data.frame} as input as well as return value. The only responsibility of that function is to draw a sample, which makes it easy to find, understand and reuse when published. The operator \proglang{\%>\%} is used to add new components to the setup.

\subsection{Model-based simulation}
\label{sec:csModel}
In the following we want to show how to construct a simulation in a model-based setting. The goal is to estimate the domain predictions under a FH model. Components which are involved are \textit{data generation} and \textit{computing on aggregated data} (see figure \ref{fig:flowdiagram}). The first step is to generate the data under the model:

\[ y_i = 100 + 2 \cdot x_i + v_i + e_i\]

where $x_i \stackrel{iid}{\sim} N(0, 4^2)$, $v_i \stackrel{iid}{\sim} N(0, 1)$ and $e_i \stackrel{indep}{\sim} N(0, \sigma_i^2)$ with $\sigma_i^2 = 0.1, 0.2, \dots, 4$ and $i = 1, \dots, 40$ as index for the domains. Also are $x_i$, $v_i$ and $e_i$ independent from each other. The area-level data for the simulation is generated in each Monte Carlo repetition.

In this case the \textit{base-component} is a data table with an id variable named \proglang{idD} and constructed with the function \proglang{base\_id}. Any random number generator in \proglang{R} can be used, however we have normally distributed variates, for which some predefined functions are available.

@@

In the design-based simulation we want to illustrate the use of \proglang{saeSim} when starting from data of a population. For this purpose we use a synthetic population generated from Austrian EU-SILC (European Union Statistics on Income and Living Conditions) data. The data consists of 25 thousand households and does not represent the true population size of Austria. It is published alongside the \proglang{R}-package \proglang{simFrame} \citep{Alf10} where it is used as an example data set. To keep this study as simple as possible, we further restrict the data for the main income holder and will only use some of the provided auxiliary information.

@@

Using this data as population, we want to repeatedly draw samples from it, then predict the domain means by using a direct estimator and a unit-level model. The sampling design is to draw 10 per cent from each region with simple random sampling. For each region the direct estimator for income as well as the EBLUP under the BHF model is computed. Although the data offers some more information, we use only \proglang{gender}, \proglang{age} and \proglang{agesq} as predictor variables. The function \proglang{eblupBHF} from the package \proglang{sae} is an implementation of the BHF estimator and used as an example. This function expects three data objects and returns domain predictions. The sampled data, the population means of the auxiliary variables and the population sizes in each domain. Before we can begin to construct the simulation setup we store these data tables as attributes to the population data. There are other options but setting attributes to the processed data is a very flexible form of processing data on different aggregation levels.

@@

What now needs to be done, is to define the components which add the desired estimates to the data. Here we will compute the mean of income in each domain as the direct estimator and the EBLUP under the BHF model. Although this could be done in one step, we will separate the two computations to illustrate how to combine several estimations and define each component independent of one another. This focus on the definition of each component and meeting the convention of the defined interface is the intended approach. It will automatically organise the simulation and each component is arranged using the simulation framework. Hence, we will define two functions, one for adding the direct estimates and one for adding the EBLUP.

@@

Another positive aspect of the above definitions is, that each step is relatively small and the purpose clearly defined, which will make them easy to understand and reuse. Finally the simulation results are combined in an \textit{aggregation-component}, possibly followed by the application of area-level models but omitted in this example. The result of this aggregation step will be a \proglang{data.frame} with one row for each region.

@@

To repeat the simulation $R = 50$ times the simulation setup can be passed to the function \proglang{sim}, the resulting \proglang{list} is directly combined using \proglang{rbind\_all}.

@@

To further process the simulation results we present two plots using the package \proglang{ggplot2} \citep{wickham09} and \proglang{reshape2} \citep{wickham07} for further reshaping of the data. As can be seen, little effort is needed to process the result data. Figure \ref{fig:dbBIAS} and \ref{fig:dbMSE} show the Monte-Carlo BIAS and MSE for each region in Austria and for each estimator. Keep in mind that the BHF was applied to illustrate the simulation framework. There are a number of issues with regard to model choice and variable selection which we will not address any further.

@@

\begin{figure}[!h]
\includegraphics[width = \textwidth]{saeSim-designSimMSE.pdf}
\caption{Monte Carlo MSE of direct vs. BHF predictor. 50 predictions for each region and estimation technique.}
\label{fig:dbMSE}
\end{figure}

\section{Outlook}
\label{sec:outlook}
With \proglang{saeSim} we have three objectives, making simulation tools available and reusable, unify the process behind simulations and third assist the researcher to design her code in a transparent way. By providing the package we not only want to make our own tools available but also open it for contributions. The package source is available on CRAN (\url{http://cran.r-project.org/web/packages/saeSim/}) and the repository for development on GitHub (\url{https://github.com/wahani/saeSim}). As GitHub allows to share and contribute source code using version-control, it is open for submissions. Apart from the availability of specific utility functions, we hope to promote and support the design of source code for simulation studies. One aspect is the design of simulations as processes of data. Furthermore, we encourage the definition of small and self contained components, i.e. functions, as this reduces the lines of code necessary to be read in order to understand it's purpose.

The package already provides more than introduced in this article. To mention is the support of outlier contaminated data. At this time only representative outliers (outlying observations in the population) are supported \citep[cf.][]{Cha86}. However, we plan to extend this feature to non-representative outliers (outliers are part of the sample but not the population). Furthermore, the inclusion of available random number generators available in the R-language is already possible and access to generate group effects as needed in mixed models is supported. Also, as the response is created by an \proglang{R} expression, any form of non-linearity in the relationship between response and auxiliary variables as well as error components can be modelled.

A more technical feature is a back-end for parallel computations which is a link to the \proglang{parallel} package in \proglang{R} \citep{r14}. Tools to process result data after the simulation as well as plotting methods for that data are an avenue for further research. Already available are some simple plots for the simulation setups as well as a summary method to get information on the expected runtime and structure of the resulting data.

>>>>

Without knowing anything about the setup defined in \proglang{sim\_base\_lm} we notice that \proglang{setup1} and \proglang{setup2} only differ in the applied sampling scheme. \proglang{sim\_sample} is responsible to control when a function is applied (after the population-level) and \proglang{sample\_number(5)} and \proglang{sample\_fraction(0.05)} define the explicit way of drawing samples. \red{Separating} the responsibility of each component into what is applied and when it is applied makes it possible to add new components to any step in the process. The composition of a simulation in that manner will focus on the definition of components and hide control structures. Any function can be passed to \proglang{sim\_sample} which has a \proglang{data.frame} as input as well as return value. The only responsibility of that function is to draw a sample, which makes it easy to find, understand and reuse when published. The operator \proglang{\%>\%} is used to add new components to the setup.

\subsection{Model-based simulation}
\label{sec:csModel}
In the following we show one way how to construct a simulation in a model-based setting. The aim is to estimate the domain predictions under a FH model. Involved components are \textit{data generation} and \textit{computing on aggregated data} (cf. Figure \ref{fig:flowdiagram}). The first step is to generate the data under the model:

\[ y_i = 100 + 2 \cdot x_i + v_i + e_i,\]

where $x_i \stackrel{iid}{\sim} N(0, 4^2)$, $v_i \stackrel{iid}{\sim} N(0, 1)$ and $e_i \stackrel{indep}{\sim} N(0, \sigma_i^2)$ with $\sigma_i^2 = 0.1, 0.2, \dots, 4$ and $i = 1, \dots, 40$ as index for the domains. $x_i$, $v_i$ and $e_i$ are independent from each other. The area-level data for the simulation is generated in each Monte Carlo replication.

In this case the \textit{base-component} is a data table with an id variable named \proglang{idD} and constructed with the function \proglang{base\_id}. Any random number generator in \proglang{R} can be used. However, we have normally distributed variates, for which some predefined functions are available in the package.

@@

In the design-based simulation we illustrate the use of \proglang{saeSim} under a fixed population. For this purpose we use a synthetic population generated from Austrian EU-SILC (European Union Statistics on Income and Living Conditions) data. The data consists of 25 thousand households. It is published alongside the \proglang{R}-package \proglang{simFrame} \citep{Alf10} where it is also used as an example data set. To keep this study as simple as possible, we further restrict the data on the main income holder and only use some of the provided auxiliary information.

@@

Using this data as population, we repeatedly draw samples from it. Then we predict the domain means by using a direct estimator and a unit-level model. The sampling design is to draw a 10 per cent sample from each region with simple random sampling. For each region the direct estimator for income and the EBLUP under the BHF model is computed. Although the data offers some more information, we only use \proglang{gender}, \proglang{age} and \proglang{agesq} as covariates. The function \proglang{eblupBHF} from the package \proglang{sae} is an implementation of the BHF estimator. This function expects three data objects and returns domain predictions. The three objects are the sampled data, the population means of the auxiliary variables and the population sizes in each domain. Before we begin to construct the simulation setup, we store these data tables as attributes to the population data. There are other options but \red{adding attributes to the data} is a very flexible form of processing data on different aggregation levels.

@@

In the next step, we define components which \red{add the estimates of interest} to the data. Here we compute the direct estimator of the mean income in each domain and the EBLUP under the BHF model. Although this could be done in one step, we separate the two computations to illustrate how to combine several estimations and how to define each component independent of each other. This focus on the definition of each component and  on meeting the convention of the defined interface is the intended approach. It automatically organises the simulation and each component is arranged using the simulation framework. Hence, we define two functions, one for adding the direct estimates and one for adding the EBLUP.

@@

\red{A positive} aspect of the above definitions is, that the code for each step is relatively short and the purpose is clearly defined. This may help to improve readability and to reproduce the research. Finally, the simulation results are combined in an \textit{aggregation-component}, possibly followed by the application of area-level models. The result of this aggregation step is a \proglang{data.frame} with one row for each region.

@@

To repeat the simulation $R = 50$ times the simulation setup is passed to the function \proglang{sim}. The resulting \proglang{list} is directly combined using \proglang{rbind\_all}.

@@

To further process the simulation results we present two plots using the package \proglang{ggplot2} \citep{wickham09} and \proglang{reshape2} \citep{wickham07} for further reshaping of the data. Figure \ref{fig:dbBIAS} and \ref{fig:dbMSE} show the Monte-Carlo BIAS and MSE for each region in Austria and for each estimator. Keep in mind that the BHF was applied to illustrate the simulation framework. There are a number of issues with regard to model choice, variable selection and outliers, which we will not discuss in this context.

@@

\begin{figure}[!h]
\includegraphics[width = \textwidth]{saeSim-designSimMSE.pdf}
\caption{Monte Carlo MSE of direct vs. BHF predictor. 50 predictions for each region and estimation technique.}
\label{fig:dbMSE}
\end{figure}
\clearpage

\section{Outlook}
\label{sec:outlook}
From our perspective, \red{there is a the need} for sharing tools for data generation and simulation amongst the scientific community to guarantee the reproducibility of research. \red{\proglang{saeSim} may provide an adequate framework for pursuing this aim in the field of small area estimation. By defining the steps of a simulation we may promote a reasonable way to communicate results in academic articles and during research.}

The package source is available on CRAN (\url{http://cran.r-project.org/web/packages/saeSim/}) and the repository for development on GitHub (\url{https://github.com/wahani/saeSim}). As GitHub allows to share and contribute source code using version-control, it is open for submissions. Apart from the availability of specific utility functions, we may promote and support the design of source code for simulation studies. One aspect is the design of simulations as processes of data. Furthermore, we encourage the definition of small and self contained components, i.e. functions. This reduces the lines of code necessary to be read in order to understand it's purpose.

The package provides more features than introduced in this article. To mention is the support of outlier contaminated data. At this time only representative outliers (outlying observations in the population) are supported \citep[cf.][]{Cha86}. However, we plan to extend this feature to non-representative outliers (outliers are part of the sample but not the population). \red{Furthermore, an interface to random number generators available in the R-language is available.} The user can also generate group effects as needed in mixed models. As the response is created by an \proglang{R} expression, any form of non-linearity in the relationship between response and auxiliary variables as well as error components can be modelled.

A more technical feature is a back-end for parallel computations which is a link to the \proglang{parallel} package in \proglang{R} \citep{r14}. \red{Tools to process result data after the simulation, i.e. summaries as well as plotting methods, are an avenue for further research.} Already available are some simple plots for the simulation setups as well as a summary method to get information on the expected runtime and structure of the resulting data.
